Acid-induced protein gels: from gelation to stress-induced failure

### 2016-06-08 creep_beamer.tex
Strain in appendix Gallot 2013 The simplest yoghurt
Adapt to iLM template

\node[shape=cylinder, draw, shape border rotate=90,minimum width=\mydimb, minimum height=7\baselineskip, anchor=north, aspect=2] at (rheometer.north-|webcam.north) (Stator) {};

@@

\begin{frame}{The yoghurt creep experiment}
\let\mylena\relax
\newlength\mylena
\pgfmathsetlength{\mylena}{0.19\textwidth}
\includegraphics[width=\mylena]{Y110_2013-03-01_02-55-11.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-06-28.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-19-47.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-10.jpg}\hfill
%\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-18.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-19.jpg}\\
%\begin{center}

@@

\begin{frame}{Different (constant) stresses: universality}
\begin{columns}
\column{0.6\textwidth}
\tikzsetnextfilename{gdots_not_normed}%
\begin{tikzpicture}
\begin{loglogaxis}[
	name=g,
	width=\textwidth,
	height=10\baselineskip,
	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
	ylabel={Strain rate $\dot{\gamma}$ (\si{\per\second})},
	cycle list name=earthy,
	no marks,
	]
% 	\fill[Accent1!20]
% 		(axis cs:10,0.4) ellipse[rotate=-20, x radius=0.15\textwidth, y radius=0.02\textwidth]
% 		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
% 		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
% 		;
	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
	%\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[sloped, pos=0.75, below left] {Primary creep};

\end{loglogaxis}
\begin{scope}[anchor=base east, every node/.style={ rotate=90}]
	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
	\node[red!60!black] at (s300 |- g.outer north) {300};
	\node[red!80!black] at (s400 |- g.outer north) {400};
	\node[red] at (s550 |- g.outer north) {550};
	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
\end{scope}
\end{tikzpicture}

\column{0.4\textwidth}
\structure{Normalisation}
\begin{description}[$\dot\gamma_\text{min}$]
\item[$\tau_f$] failure time 
\item[$\dot\gamma_\mathrm{min}$] minimum strain rate 
\end{description}

\smallskip
\hfill$\Rightarrow$ \colorbox{Accent1}{\color{white}\Large Physical origin?}
\end{columns}

@@

\begin{frame}{Power-law creep and plasticity}
\begin{block}{``Andrade'' creep (1910)}
\begin{itemize}
\item Power-law creep known in \alert{ductile solids} for over a century
\item Classically explained by disinclination dynamics in crystals
\item Recently explained by plastic events in glasses \textit{\footnotesize Miguel et al., PRL 2002}
\end{itemize}
\end{block}

@@

\bigskip
\movie[externalviewer]{\structure{Ultrasound velocimetry}}{Yaourt110_primary.avi} $\Rightarrow$ local velocity maps
\begin{itemize}
	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
	\item if present, plastic events are below our resolution (a few \si{\micro\metre})
\end{itemize}
\end{frame}

\begin{frame}{Is power-law creep damaging the structure?}
\begin{itemize}
\item Stop creep much before ($\gamma<0.3$) fracture nucleation ($\gamma\approx 1$)
\item Let relax $\Rightarrow$ No complete recovery
\item Reset deformation and start again $\Rightarrow$  superimposable
\end{itemize}

@@

\begin{itemize}
\item Undamaged structure\raisebox{-0.6\baselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.\baselineskip}\right\}\Rightarrow\text{Plasticity cannot explain power law}$}
\item Viscous dissipation
\end{itemize}
\end{frame}

@@

\begin{frame}{Fibre bundle models}
	\tikzsetnextfilename{FBM_Jagla}
	\begin{tikzpicture}
	\let\mydima\relax
	\newlength\mydima
	\setlength{\mydima}{0.3\textwidth}
	\let\mydimb\relax
	\newlength\mydimb
	\setlength{\mydimb}{2\baselineskip}

@@

	\node[above=0 of top, single arrow, draw, shape border rotate=90] {$\sigma$};
	\node[starburst, draw=Accent2, inner sep=0pt,  minimum width=0.1\mydima, minimum height=0.5\mydimb, starburst point height=0.15em] at (-0.1\mydima,0) {};
	\node[double arrow,draw, anchor=west, text width=0.21\textwidth, align=center] at (0.5\mydima,-0.5\mydimb) (model) {\structure{model}};
	\node[above left=\baselineskip and 0of model.east, text width=0.31\textwidth, inner sep=0] {\begin{itemize}
		\item elastic fibres
		\item local yield strain
		\item local coupling
		\end{itemize}};
	\end{tikzpicture}
	\includegraphics[height=6\baselineskip, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}

@@

	\column{0.3\textwidth}
	\structure{But}
	\begin{itemize}
	\item yield stress
	\item built in power-law creep
	\item no master curve
	\end{itemize}

	\column{0.4\textwidth}\tikzsetnextfilename{creep_1_3}
	\begin{tikzpicture}
		\begin{groupplot}[%
			group style={
				group name=g, group size=1 by 2,
				vertical sep=2.5em,
				},
			height=6.5\baselineskip,
			width=9\baselineskip,
			cycle list name=earthy,
			clip mode=individual,
			ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
			xlabel shift=-0.5em,
			]
		\nextgroupplot[
			xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
			xmode=log,ymode=log,
			xtick={1e-4, 1e-2, 1},
			]
			\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
			\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};

@@

\begin{frame}{Fibre bundle models}
	\begin{columns}[T]
	\column{0.27\textwidth}
	\textit{\footnotesize Kun et al.  PRL 2008}

@@

	\textit{\footnotesize Halász et al., PRE 2012}

@@

	\tikzsetnextfilename{model_arrow}
	\begin{tikzpicture}
	\node[double arrow,draw, anchor=west, text width=\textwidth-1\baselineskip, align=center] (model) {\structure{model}};
	\end{tikzpicture}

@@

	\column{0.38\textwidth}
	\tikzsetnextfilename{basquin_fracture}
	\begin{tikzpicture}
	\begin{loglogaxis}[
		name=a,
		width=\textwidth,
		height=0.8\textwidth,
		xlabel={$\sigma$ (\si{\pascal})},
		ylabel={$\tau_f$ (\si{\second})},
		xmin=100, ymin=10, ymax=1e6,
		xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
		]
		\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
		\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)};
	\end{loglogaxis}
	\node[below = of a]{\includegraphics[height=6\baselineskip]{Y110_2013-03-01_03-21-15.jpg}};
	\end{tikzpicture}

@@

\begin{frame}{Creep and yielding summary}
\begin{itemize}
\item failure involves a single time scale $\tau_f \sim \sigma^{-5.5}$ Basquin law
\item reversible, homogeneous creep $\rightarrow$ irreversible fracture growth
\item a single expression captures all the global rheological response
\end{itemize}
\tikzsetnextfilename{creep_summary}
\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=3 by 1,
			horizontal sep=3em,
			},
		width=0.35\textwidth,
		cycle list name=earthy,
		clip mode=individual,
		]
	\nextgroupplot[
		xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
		ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
		xmode=log,ymode=log,
		xtick={1e-4, 1e-2, 1},
		]
		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
		\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};

>>>>

\node[shape=cylinder, draw, shape border rotate=90,minimum width=\mydimb, minimum height=7\baselineskip, anchor=north, aspect=2, yshift=-1.5em] at (rheometer.north-|webcam.north) (Stator) {};

@@

\begin{frame}{The yoghurt creep experiment}
\vspace{1.25\baselineskip}
\let\mylena\relax
\newlength\mylena
\pgfmathsetlength{\mylena}{0.19\textwidth}
\includegraphics[width=\mylena]{Y110_2013-03-01_02-55-11.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-06-28.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-19-47.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-10.jpg}\hfill
%\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-18.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-19.jpg}\\
%\begin{center}

@@

\begin{frame}{Different stresses: universality}
\begin{columns}
\column{0.4\textwidth}
\textit{\footnotesize constant stresses}\\
\vspace{\baselineskip}
\structure{Normalisation}
\begin{description}[$\dot\gamma_\text{min}$]
\item[$\tau_f$] failure time 
\item[$\dot\gamma_\mathrm{min}$] minimum strain rate 
\end{description}

\smallskip
\hfill\begin{beamercolorbox}[rounded=true]{block title}\Large $\Rightarrow$Physical origin?\end{beamercolorbox}

\column{0.6\textwidth}
\tikzsetnextfilename{gdots_not_normed}%
\begin{tikzpicture}
\begin{loglogaxis}[
	name=g,
	width=\textwidth,
	height=10\baselineskip,
	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
	ylabel={Strain rate $\dot{\gamma}$ (\si{\per\second})},
	cycle list name=earthy,
	no marks,
	]
% 	\fill[Accent1!20]
% 		(axis cs:10,0.4) ellipse[rotate=-20, x radius=0.15\textwidth, y radius=0.02\textwidth]
% 		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
% 		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
% 		;
	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
	%\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[sloped, pos=0.75, below left] {Primary creep};

\end{loglogaxis}
\begin{scope}[anchor=base east, every node/.style={ rotate=90}]
	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
	\node[red!60!black] at (s300 |- g.outer north) {300};
	\node[red!80!black] at (s400 |- g.outer north) {400};
	\node[red] at (s550 |- g.outer north) {550};
	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
\end{scope}
\end{tikzpicture}
\end{columns}

@@

\begin{frame}{Power-law creep and plasticity}
\begin{block}{``Andrade'' creep (1910)}
\begin{itemize}
\item Power-law creep known in \alert{ductile solids} for over a century
\item Classically explained by disinclination dynamics in crystals
\item Recently explained by plastic events in glasses\\\hfill \textit{\footnotesize Miguel et al., PRL 2002}
\end{itemize}
\end{block}

@@

\bigskip
\movie[externalviewer]{\structure{Ultrasound velocimetry}}{Yaourt110_primary.avi} $\Rightarrow$ local velocity maps\\\hfill\textit{\footnotesize T. Gallot et al., Rev. Sci. Instrum. 84, 045107 (2013)}
\begin{itemize}
	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
	\item if present, plastic events are below our resolution {\small (a~few~\si{\micro\metre})}
\end{itemize}
\end{frame}

\begin{frame}{Is creep damaging the structure?}
\vspace{0.25\baselineskip}
\begin{itemize}
\item Stop creep much before ($\gamma<0.3$) fracture nucleation ($\gamma\approx 1$)
\item Let relax $\Rightarrow$ No complete recovery
\item Reset deformation and start again $\Rightarrow$  superimposable
\end{itemize}

@@

\begin{itemize}
\item Undamaged structure\raisebox{-0.6\baselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.\baselineskip}\right\}\Rightarrow\begin{array}{c}\text{Plasticity cannot explain}\\ \text{power law}\end{array}$}
\item Viscous dissipation
\end{itemize}
\end{frame}

@@

\begin{frame}{Fibre bundle models}
	\vspace{\baselineskip}
	\tikzsetnextfilename{FBM_Jagla}%
	\begin{tikzpicture}
	\let\mydima\relax
	\newlength\mydima
	\setlength{\mydima}{0.3\textwidth}
	\let\mydimb\relax
	\newlength\mydimb
	\setlength{\mydimb}{2\baselineskip}

@@

	\node[above=0 of top, single arrow, draw, shape border rotate=90] {$\sigma$};
	\node[starburst, draw=Accent2, inner sep=0pt,  minimum width=0.1\mydima, minimum height=0.5\mydimb, starburst point height=0.15em] at (-0.1\mydima,0) {};
		\node[double arrow,draw, anchor=west, text width=0.25\textwidth, align=center] at (0.5\mydima,-0.5\mydimb) (model) {\structure{model}};
	\node[above left=\baselineskip and 0of model.east, text width=0.35\textwidth, inner sep=0] {\begin{itemize}
		\item elastic fibres
		\item local yield strain
		\item local coupling
		\end{itemize}};
	\end{tikzpicture}
	\includegraphics[width=0.35\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}

@@

	\column{0.34\textwidth}
	\structure{But}
	\begin{itemize}
	\item yield stress
	\item built in power-law creep
	\item no master curve
	\end{itemize}

	\column{0.4\textwidth}\tikzsetnextfilename{creep_1_3}%
	\begin{tikzpicture}
		\begin{groupplot}[%
			group style={
				group name=g, group size=1 by 2,
				vertical sep=2.5em,
				},
			height=6.5\baselineskip,
			width=9\baselineskip,
			cycle list name=earthy,
			clip mode=individual,
			ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
			xlabel shift=-0.5em,
			]
		\nextgroupplot[
			xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
			xmode=log,ymode=log,
			xtick={1e-4, 1e-2, 1},
			]
			\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
			\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
			\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};

@@

\begin{frame}{Fibre bundle models}
	\begin{columns}
	\column{0.27\textwidth}
	\vspace{0.5\baselineskip}
	\textit{\scriptsize Kun et al.  PRL 2008}

@@

	\textit{\scriptsize Halász et al. PRE 2012}

@@

	\tikzsetnextfilename{model_arrow}%
	\begin{tikzpicture}
	\node[double arrow,draw, anchor=west, text width=\textwidth-1\baselineskip, align=center] (model) {\structure{model}};
	\end{tikzpicture}

@@

	\column{0.38\textwidth}
	\tikzsetnextfilename{basquin_fracture}%
	\begin{tikzpicture}
	\begin{loglogaxis}[
		name=a,
		width=\textwidth,
		height=0.8\textwidth,
		xlabel={$\sigma$ (\si{\pascal})},
		ylabel={$\tau_f$ (\si{\second})},
		xmin=100, ymin=10, ymax=1e6,
		xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
		]
		\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
		\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)};
	\end{loglogaxis}
	\node[below = of a]{\includegraphics[height=6\baselineskip]{Y110_2013-03-01_03-21-15.jpg}};
	\end{tikzpicture}

@@

\begin{frame}{Creep and yielding summary}
\begin{itemize}
\item failure involves a single time scale $\tau_f \sim \sigma^{-5.5}$ Basquin law
\item \begin{minipage}{0.4\textwidth}\center reversible,\\ homogeneous creep\end{minipage} \hfill\textcolor{ilmorange}{$\rightarrow$}\hfill \begin{minipage}{0.4\textwidth}\center irreversible fracture growth\end{minipage}
\item a single expression captures all the global rheological response
\end{itemize}
\tikzsetnextfilename{creep_summary}
\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=3 by 1,
			horizontal sep=2.5em,
			},
		width=0.375\textwidth,
		cycle list name=earthy,
		clip mode=individual,
		]
	\nextgroupplot[
		xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
		ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
		xmode=log,ymode=log,
		xtick={1e-4, 1e-2, 1},
		]
		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
		\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};
